Service Oriented ARC Storage; \\An Overview and Design

### 2008-12-05 doc/storage_system/ARCStorage_src/ARCStorage.tex
Some bibliography and typo fixes, added support text to acknowledgements and made addresses real

\begin{document}
\pagestyle{headings} 
\title{Service Oriented ARC Storage; \\An Overview and Design}
%% \author{  Jon K. Nilsen$^{2,3}$ \and Salman Toor$^{1}$
%%   \and Zsombor Nagy$^{5}$ \and Bjarte Mohn$^{4}$  \\ $^{1}$Scientific
%%   Computing Division, IT Department, Uppsala University, Sweden
%%   \\ $^{2}$Center for Information Technology, University of Oslo, Norway
%%   \\ $^{3}$Department of Physics,  University of Oslo, Norway
%%   \\  $^{4}$Department of Physics and Astronomy, Uppsala University,
%%   Sweden \\ $^{5}$National Information Infrastructure Development
%%   Institute, Hungary}
\author{  Jon K. Nilsen$^{\dagger,\ddagger}$ \and Salman Toor$^{*}$
  \and Zsombor Nagy$^{\mathparagraph}$ \and Bjarte Mohn$^{\mathsection}$  \\ $^{*}$Scientific
  Computing Division, IT Department, Uppsala University, Sweden
  \\ $^{\dagger}$Center for Information Technology, University of Oslo, Norway
  \\ $^{\ddagger}$Department of Physics,  University of Oslo, Norway
  \\  $^{\mathsection}$Department of Physics and Astronomy, Uppsala University,
  Sweden \\ $^{\mathparagraph}$National Information Infrastructure Development
  Institute, Hungary\\ \\  j.k.nilsen@usit.uio.no \hspace{5 mm}
  salman.toor@it.uu.se \hspace{5 mm} zsombor@niif.hu \\  \hspace{7 mm}
  bjarte.mohn@fysast.uu.se}

@@

The challenge of building a reliable, self-healing, fault-tolerant,
consistent data management system in a web scale is an interesting
task. Making the system work in a heterogeneous, distributed
environment like the Grid is even more interesting. An increasing
number of applications demand not only increased CPU power, but also
vast amounts of storage space. The required storage space is not only
restricted to the duration in which the application runs; the data
should often be available for years afterwards, in certain cases even
for decades. Nowadays, we can easily find single Grid jobs which produce
gigabytes or even terabytes of data, ramping up the requirements of
storage systems to the petabyte scale and beyond. Hard drive capacity is
still increasing and storage raids can handle failing hard drives through 
data replication. However, even a designated storage site cannot guarantee 
24/7 availability or long term security against catastrophic losses of data. 
The need for a distributed, self-healing storage system is evident. To
make the storage system useable to Grid users, the system must provide
reliable and secure file transfer protocols, a cataloging system and
secure storage of the data. Several projects and designs have emerged
to address such challenges \cite{Hoschek00datamanagement,DengWang}.

In the advent of the next generation Advanced Resource Connector (ARC)
Grid middleware \cite{NorduGridsite} (due during Fall 2009), we present 
the ARC storage system \cite{ARCStoragedesigndoc}. This distributed
storage system is designed to provide an easy
to use, flexible and scalable system that can offer native storage and
at the same time provide access to third-party solutions like dCache \cite{dCache,DSSWithdCache}  by
using the same, uniform interface. Being part of ARC, the
storage system is based on a service
oriented architecture, in which each major component of the
system runs as a separate service within the ARC Hosting Environment
Daemon (HED) \cite{HEDdesigndoc}. The HED service container gives the
capability of flexible replacement of the components as well as the
possibility to introduce modifications in the future. %% The HED service container gives
%% the capability of flexible replacement of the components as well as
%% the future modification/enhancement of the system.

@@

\section{The Advance Resource Connector}
\label{The Advance Resource Connector}

The next generation Advanced Resource Connector (ARC) Grid middleware 
\cite{NorduGridsite} \cite{KnowARCsite} consists of a set of pluggable 
components. These components are the
fundamental building blocks of the ARC services and clients. ARC
services run inside a container called the Hosting Environment
Daemon (HED) and there are four kinds of pluggable components with well
defined tasks: Data Management Components are used to transfer the
data using various protocols, Message Chain Components are
responsible for the communication within clients and services as well as
between the clients and the services, ARC Client Components are
plug-ins used by the clients to connect to different Grid flavors, and
Policy Decision Components are responsible for the security model
within the system.

@@

\begin{itemize} 
\item Each  file and collection have a unique ID called GUID.
\item Services are uniquely identified by a serviceID. 
\item The Shepherds in the system identify their files by a referenceID.
\end{itemize}

@@

Replicas in a distributed storage system can have different states;
they can be broken, deleted, partially uploaded, etc. In the
ARC Storage, all replicas have assigned a state, some of which are
`\textbf{alive}' (if the replica passed the checksum test, and the
Shepherd reports that the storage node is healthy), `\textbf{invalid}'
(if the replica has wrong checksum, or the Shepherd claims it has no
such replica), `\textbf{creating}' (if the replica is in
the state of uploading).

@@

\subsection{Core Components} 
\subsubsection{Bartender}
\label{Bartender}
The Bartender service provides a high-level interface for the storage
system. Clients connect to the Bartender to create and remove
files, collections and mount-points using their Logical Names. The
Bartender communicates 
with the Librarian and Shepherd services to execute the clients'
requests. However, the actual file data does not go through the
Bartender, instead file
transfers are directly performed between the storage nodes and the
clients. There could be any number of independent Bartender services
running in the system, providing high-availability and
load-balancing. 
\subsubsection{Librarian}
\label{librarian}
The Librarian manages the hierarchy and metadata of
files, collections and mount points, as well as the health information of the Shepherd
services. In addition, the Librarian handles
the information about registered Shepherd services. The Librarian receives
heartbeat messages from the Shepherds and change replica states automatically
if needed. The Librarian uses the A-Hash for consistently storing all metadata. This makes the
Librarian a stateless service, thus enabling the system to have any
number of independent
Librarian services, again providing high-availability in the
system.

@@

The services were distributed as shown in Fig. \ref{fig:map}:
\begin{itemize}
\item An A-Hash runs at UPPMAX.
\item A Bartender runs at USIT.
\item A Librarian runs on UPPMAX.
\item In total five Shepherds were used for the tests: Three at USIT, having
  100GB storage space each, one at UPPMAX, with 20GB, and one at NIIF, providing 16GB of storage space.
 \end{itemize}

@@

Some key areas need special attention and effort to make the proposed
storage system even more stable, reliable and consistent:
\begin{itemize}
\item \textbf{Security:} This is still under development. The design is more
  or less ready, but it still needs to be implemented and properly
  tested.
\item \textbf{The A-Hash}: This is currently centralized. To avoid a singel
  point of failure in the system, and to improve the system
  performance, the A-Hash needs to be distributed.
\item \textbf{Performance optimization:} To make a storage system
  ready for production, one needs to discover and improve on
  possible bottlenecks within the system. As soon as the A-Hash is
  distributed, other possible bottlenecks, such as load-balancing
  and self-healing mechanisms, will be investigated.
\item \textbf{Protocols:} To ease the interoperability with third-party
  storage solutions and clients, the system needs to support storage
  protocols such as SRM and GridFTP. In addition the system will come
  with its own ARC protocol.
\end{itemize}

@@

>>>>

\begin{document}
\pagestyle{headings} 
\title{Service Oriented ARC Storage; \\An Overview and Design}
%% \author{  Jon K. Nilsen$^{2,3}$ \and Salman Toor$^{1}$
%%   \and Zsombor Nagy$^{5}$ \and Bjarte Mohn$^{4}$  \\ $^{1}$Scientific
%%   Computing Division, IT Department, Uppsala University, Sweden
%%   \\ $^{2}$Center for Information Technology, University of Oslo, Norway
%%   \\ $^{3}$Department of Physics,  University of Oslo, Norway
%%   \\  $^{4}$Department of Physics and Astronomy, Uppsala University,
%%   Sweden \\ $^{5}$National Information Infrastructure Development
%%   Institute, Hungary}
\author{  Jon K. Nilsen$^{\dagger,\ddagger}$ \and Salman Toor$^{*}$
  \and Zsombor Nagy$^{\mathparagraph}$ \and Bjarte Mohn$^{\mathsection}$ 
\\$^{*}$Dept. of Information Technology, Div. of Scientific Computing, Uppsala
University,\\ Box~337, SE-75105~Uppsala, Sweden
 \\ $^{\dagger}$University of Oslo, Center for Information Technology,\\
P.~O.~Box~1059, Blindern, N-0316~Oslo, Norway
  \\ $^{\ddagger}$University of Oslo, Dept. of Physics, P.~O.~Box~1048,
Blindern, N-0316~Oslo, Norway
  \\  $^{\mathsection}$Dept. of Physics and Astronomy, Div. of Nuclear and
Particle Physics,\\ Uppsala University, Box~535, SE-75121~Uppsala, Sweden \\
$^{\mathparagraph}$Institute of National Information and Infrastructure
Development\\ NIIF/HUNGARNET, Victor~Hugo~18-22, H-1132~Budapest, Hungary\\ \\ 
j.k.nilsen@usit.uio.no \hspace{5 mm}
  salman.toor@it.uu.se \hspace{5 mm} zsombor@niif.hu \\  \hspace{7 mm}
  bjarte.mohn@fysast.uu.se}

@@

The challenge of building a reliable, self-healing, fault-tolerant,
consistent data management system in a web scale is an interesting
task. Making the system work in a heterogeneous, distributed
environment like the Grid is even more interesting. An increasing
number of applications demand not only increased CPU power, but also
vast amounts of storage space. The required storage space is not only
restricted to the duration in which the application runs; the data
should often be available for years afterwards, in certain cases even
for decades. Nowadays, we can easily find single Grid jobs which produce
gigabytes or even terabytes of data, ramping up the requirements of
storage systems to the petabyte scale and beyond. Hard drive capacity is
still increasing and storage raids can handle failing hard drives through 
data replication. However, even a designated storage site cannot guarantee 
24/7 availability or long term security against catastrophic losses of data. 
The need for a distributed, self-healing storage system is evident. To
make the storage system useable to Grid users, the system must provide
reliable and secure file transfer protocols, a cataloging system and
secure storage of the data. Several projects and designs have emerged
to address such challenges~\cite{Hoschek00datamanagement,DengWang}.

In the advent of the next generation of the Advanced Resource Connector (ARC)
Grid middleware~\cite{arc} (new release due during fall 2009), we present 
the ARC storage system~\cite{ARCStoragedesigndoc}. This distributed
storage system is designed to provide an easy
to use, flexible and scalable system that can offer native storage and
at the same time provide access to third-party solutions like
dCache~\cite{dCache,DSSWithdCache}  by
using the same, uniform interface. Being part of ARC, the
storage system is based on a service
oriented architecture, in which each major component of the
system runs as a separate service within the ARC Hosting Environment
Daemon (HED)~\cite{HEDdesigndoc}. The HED service container gives the
capability of flexible replacement of the components as well as the
possibility to introduce modifications in the future. %% The HED service container gives
%% the capability of flexible replacement of the components as well as
%% the future modification/enhancement of the system.

@@

\section{The Advanced Resource Connector}
\label{The Advance Resource Connector}

The next generation of Advanced Resource Connector (ARC) Grid middleware is
developed by NorduGrid~\cite{NorduGridsite} and the EU KnowARC
 project~\cite{KnowARCsite}. It consists of a set of pluggable 
components. These components are the
fundamental building blocks of the ARC services and clients. ARC
services run inside a container called the Hosting Environment
Daemon (HED) and there are four kinds of pluggable components with well
defined tasks: Data Management Components are used to transfer the
data using various protocols, Message Chain Components are
responsible for the communication within clients and services as well as
between the clients and the services, ARC Client Components are
plug-ins used by the clients to connect to different Grid flavors, and
Policy Decision Components are responsible for the security model
within the system.

@@

\begin{itemize} 
\item Each  file and collection has a unique ID (GUID).
\item Services are uniquely identified by a serviceID. 
\item The Shepherds in the system identify their files by a referenceID.
\end{itemize}

@@

Replicas in a distributed storage system can have different states;
they can be broken, deleted, partially uploaded, etc. In the
ARC Storage, all replicas have assigned a state, some of which are
`\textbf{alive}' (if the replica passed the checksum test, and the
Shepherd reports that the storage node is healthy), `\textbf{invalid}'
(if the replica has wrong checksum, or the Shepherd claims it has no
such replica) and `\textbf{creating}' (if the replica is in
the state of uploading).

@@

\subsection{Core Components} 
\subsubsection{Bartender}
\label{Bartender}
The Bartender service provides a high-level interface for the storage
system. Clients connect to the Bartender to create and remove
files, collections and mount-points using their Logical Names. The
Bartender communicates 
with the Librarian and Shepherd services to execute the clients'
requests. However, the actual file data does not go through the
Bartender, instead file
transfers are directly performed between the storage nodes and the
clients. There could be any number of independent Bartender services
running in the system, providing high-availability and
load-balancing. 
\subsubsection{Librarian}
\label{librarian}
The Librarian manages the hierarchy and metadata of
files, collections and mount points, as well as the health information of the Shepherd
services. In addition, the Librarian handles
the information about registered Shepherd services. The Librarian receives
heartbeat messages from the Shepherds and changes replica states automatically
if needed. The Librarian uses the A-Hash for consistently storing all metadata. This makes the
Librarian a stateless service, thus enabling the system to have any
number of independent
Librarian services, again providing high-availability in the
system.

@@

The services were distributed as shown in Fig. \ref{fig:map}:
\begin{itemize}
\item An A-Hash runs at UPPMAX.
\item A Bartender runs at USIT.
\item A Librarian runs at UPPMAX.
\item In total five Shepherds were used for the tests: Three at USIT, having
  100GB storage space each, one at UPPMAX, with 20GB, and one at NIIF, providing 16GB of storage space.
 \end{itemize}

@@

Some key areas need special attention and effort to make the proposed
storage system even more stable, reliable and consistent:
\begin{itemize}
\item \textbf{Security:} This is still under development. The design is more
  or less ready, but it still needs to be implemented and properly
  tested.
\item \textbf{The A-Hash}: This is currently centralized. To avoid a single
  point of failure in the system, and to improve the system
  performance, the A-Hash needs to be distributed.
\item \textbf{Performance optimization:} To make a storage system
  ready for production, one needs to discover and improve on
  possible bottlenecks within the system. As soon as the A-Hash is
  distributed, other possible bottlenecks, such as load-balancing
  and self-healing mechanisms, will be investigated.
\item \textbf{Protocols:} To ease the interoperability with third-party
  storage solutions and clients, the system needs to support storage
  protocols such as SRM and GridFTP. In addition the system will come
  with its own ARC protocol.
\end{itemize}

@@

The work has been supported by the European Commission through the KnowARC
project (contract nr. 032691) and by the Nordunet3 programme through the NGIn
project.
